---
title: "NGS (Exome) Gene Coverage Analysis"
output: 
  html_notebook:
   toc: True
---

# Load data
This is a departure from the previous installments, as we are loading in a processed dataset. The reasons why are numerous:

1. The original data is 330mb, compressed
2. After loading (2 minutes on my quite fast computer) and uncompressing, it takes over 10GB of RAM on my computer
3. The original data needed a severe amount of massaging to make it (easily) useable:
- Annotating with gene name
- Identifying primary transcript for gene
- Expanding range data into row-form

```{r}
library(tidyverse)
library(cowplot) # you probably need to install this with install.packages('cowplot')
dd_class <- read_csv('dd_class.csv')

head(dd_class)
```

# How many genes are in this dataset?
```{r}
dd_class$Name %>% unique() %>% length()
```

# What genes are in here?
```{r}
dd_class$Name %>% unique() %>% sort()
```

# How many data points (bases) per gene?
```{r}
dd_class %>% 
  group_by(Name) %>% 
  summarise(Count=n())
```

# How many exons per gene?
```{r}
dd_class %>% 
  select(Name, `Exon Number`) %>% 
  unique() %>% 
  group_by(Name) %>% 
  summarise(Count = n())
```

# How many base pairs of ABCA4 (well, ABCA4 exons) is covered by more than 10 reads?
Base R style
```{r}
# Grab the Read_Depth vector from the data frame filtered by ABCA4 values
depth_abca4 <- dd_class %>% 
  filter(Name=='ABCA4') %>% 
  pull(Read_Depth)
sum(depth_abca4 > 10)
```

# 5 reads?
```{r}
sum(depth_abca4 > 5)
```

# Let's check all of the genes to see which are the worst covered
```{r}
dd_class %>% 
  group_by(Name) %>% 
  summarise(Total_Bases = n(),
            LT5 = sum(Read_Depth < 5),
            LT10 = sum(Read_Depth < 10),
            Good = sum(Read_Depth >= 10),
            P5 = LT5 / Total_Bases,
            P10 = LT10 / Total_Bases) %>% 
  arrange(-P10)
```


# We can visually display the data, also
```{r}
dd_class %>% 
  ggplot(aes(x=Read_Depth, group=Name)) +
  geom_density()
```

# Hard to see what is going on, let's make little plots for each gene
```{r, fig.width=10, fig.height=10}
dd_class %>% 
  ggplot(aes(x=Read_Depth, group=Name)) +
  facet_wrap(~Name) + 
  geom_density()
```


# Where are genes poorly covered?
## BBIP1
```{r}
dd_class %>% filter(Name=='BBIP1') %>% 
  ggplot(aes(x=Start, y=Read_Depth)) + 
  facet_wrap(~`Exon Number`, scales = 'free_x', nrow=1, strip.position = 'bottom') + 
  geom_point(size=0.1) + theme_minimal() +
  theme(axis.text.x=element_blank(), 
        axis.ticks.x = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank(),
        legend.position = 'none') + 
  ylab('Depth') + 
  xlab('Exon Number')

```


# Make a coverage plot for many genes (This is advanced stuff!!!)

```{r, fig.width=6, fig.height=3}
gene_list <- c('ABCA4', 'PITX2','VSX2','RPGR','SOX10')
 
plot_maker <- function(gene){
  num_of_exons <- dd_class%>% filter(Name==gene) %>% pull(`Exon Number`) %>% as.numeric() %>% max()
   # expand to create a row for each sequence and fill in previous values
  dd_class %>% filter(Name==gene) %>%
    mutate(`Exon Number`= factor(`Exon Number`,levels=0:num_of_exons)) %>%  
    ggplot(aes(x=Start, y=Read_Depth)) + 
    facet_wrap(~`Exon Number`, scales = 'free_x', nrow=1, strip.position = 'bottom') + 
    geom_point(size=0.1) + theme_minimal() +
    theme(axis.text.x=element_blank(), 
          axis.ticks.x = element_blank(), 
          panel.grid.minor = element_blank(), 
          panel.grid.major.x = element_blank(),
          legend.position = 'none') + 
    ylab('Depth') + 
    xlab(gene)
}

plots <- list()
for (i in gene_list){
  plots[[i]] <- plot_maker(i)
}

plot_grid(plotlist = plots, ncol=1)
```